---
title: "Homework 2"
author: "Mengyu Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

### Read and clean the Mr. Trash Wheel sheet
* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in `read_excel`
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer and converts the result to an integer variable (using `as.integer`)
```{r read and clean Mr. Trash Wheel sheet}
df_trash = read_excel(
  path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel") %>% 
  select(-(15:17))



trash = df_trash %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))

trash
```

### Read and clean precipitation data for 2017 and 2018

For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable `month.name` is built into R and should be useful).

```{r Read and clean precipitation data for 2017 and 2018}
#import sheet 2017 precipitation
precpt_2017 = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = rep(2017))

#import sheet 2018 precipitation
precpt_2018 = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = rep(2018))

#combine two sheets
precpt_data = bind_rows(precpt_2017, precpt_2018) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, total)

precpt_data
```

### Description of Data

* Dataset trash

The number of observation is `r nrow(trash)`. It includes the `r names(trash[, 4])` of collection,  the `r names(trash[, 5])` and `r names(trash[, 6])` of the trash that each dumpter recieved, and the amount of different types of trash such as `r str_to_lower(names(df_trash[, 7:13]))`. 

* Dataset precpt_2017 and precpt_2018

The number observation in precpt_2017 and precpt_2018 are `r nrow(precpt_2017)` and `r nrow(precpt_2018)` respectively. They both includes the `r names(precpt_data[, 3])` amount of precipitation in each month between January 2017 and December 2018.

The total precipitation in 2018 is `r sum(pull(precpt_2018, total))`. 

The median number of sports balls in a dumpster in 2017 is `r median(pull(subset(trash, year == 2017), sports_balls))`.

# Problem 2

### Clean the Data in pols-month.csv
Use `separate()` to break up the variable `mon` into integer variables `year`, `month`, and `day`; replace month number with month name; create a `president` variable taking values `gop` and `dem`, and remove `prez_dem` and `prez_gop`; and remove the `day` variable.

```{r Clean the Data in pols-month.csv}
pols = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(col = mon, into = c("year", "month", "day")) %>% 
  mutate(month = as.integer(month),
         month = month.name[month],
         president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  select(-prez_dem, -prez_gop, -day)

pols
```


### Clean the Data in snp.csv

```{r Clean the Data in snp.csv}
snp = read_csv("data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c("month", "day", "year")) %>%
  select(year, month, close) %>%
  mutate(month = as.integer(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month])

snp
```

### Tidy the `unemployment` Data

This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r Tidy the Unemployment Data}
unemployment = read_csv("data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "pct_unemployment"
    ) %>%
  mutate(month = match(month,month.abb),
         month = month.name[month],
         Year = as.character(Year)) %>% 
  janitor::clean_names()
```

### Join the datasets
Join the datasets by merging snp into pols, and merging unemployment into the result. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

```{r Join the datasets}
df_join = left_join(pols, snp, by = c("year","month"))
df_join = left_join(df_join, unemployment, by = c("year","month"))
df_join
```

The dimension of `pols` is `r nrow(pols)` observations of `r ncol(pols)`. The range of `year` is `r range(pull(pols, year))`. Names of key variables are `r names(pols[, 3:9])`.

The dimension of `snp` is `r nrow(snp)` observations of `r ncol(snp)`. The range of `year` is `r range(pull(snp, year))`. Name of key variable is `r names(snp[, 3])`.

The dimension of `unemployment` is `r nrow(unemployment)` observations of `r ncol(unemployment)`. The range of `year` is `r range(pull(unemployment, year))`. Name of key variable is `r names(unemployment[, 3])`.

The dimension of `df_join` is `r nrow(df_join)` observations of `r ncol(df_join)`. The range of `year` is `r range(pull(df_join, year))`. Name of key variable is `r names(df_join[, 3:11])`.

# Problem 3
### Clean the Data
```{r}
#clean the data
baby_names = read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  rename("first_name" = childs_first_name) %>%
  mutate(first_name = str_to_lower(first_name),
         ethnicity = recode(ethnicity, 
                "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                "BLACK NON HISP" = "BLACK NON HISPANIC",
                "WHITE NON HISP" = "WHITE NON HISPANIC")) %>% 
  distinct()

```

### Show the rank in popularity of the name “Olivia” as a female baby name
```{r}
#olivia
olivia = baby_names %>% 
  filter(first_name == "olivia", gender == "FEMALE") %>% 
  select(-gender, -first_name, -count) %>% 
  pivot_wider(
    names_from = "ethnicity", 
    values_from = "rank") %>% 
  arrange(year_of_birth)

olivia
```

### The Most Popular Name among Male Children

```{r}
#the most popular name among male children

male_pop = baby_names %>% 
  filter(rank == 1, gender == "MALE") %>% 
  select(-gender, -rank, -count) %>% 
  mutate(first_name = str_to_title(first_name)) %>% 
  pivot_wider(
    names_from = "ethnicity", 
    values_from = "first_name") %>% 
  arrange(year_of_birth) 

male_pop
```

### Scatter Plot
For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
scatter_plot = baby_names %>% 
  filter(
    ethnicity == "WHITE NON HISPANIC", 
    gender == "MALE",
    year_of_birth == "2016"
    )
  
ggplot(scatter_plot, aes(x = rank, y = count)) + 
  geom_point()
```

